Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-03-23 PAPER.tex
made syntax/typo changes from full-read-through up through acknowledgements

While the null test is useful for testing noise properties and the uniformity of a dataset, jackknives are useful in pinpointing 
which data subsets are contaminated by biases and which are not; in our toy model we see that the bias exists only in $
\textbf{x}_{1}$ (Figure \ref{fig:toy_bias2}). If foreground or noise biases exist in a dataset, jackknives can tease them out and 
provide insight into possible sources. For example, if jackknives along the time-axis reveal a bias present at a certain LST, a 
likely explanation would be excess foreground emission from a radio source in the sky at that time. A jackknife test involving 
data before and after the application of a fringe-rate filter can reveal whether cross-talk noise bias is successfully suppressed 
with the filter, or if similar-shaped detections in both power spectra suggest otherwise. There are many other jackknife axes of 
which we will not go into detail here, including baseline, frequency, and polarization. Ultimately, an EoR detection should persist 
through them all and a clean measurement should exhibit noise-like null spectra.

@@

where $\nu$ are frequencies in GHz. We use a receiver temperature of $144$\,K, yielding $T_{sys} = 431$\,K at $150$\,MHz. 
This is lower than the $T_{sys}$ of $500$\,K used in \citetalias{ali_et_al2015} because of several small mis-calculation errors that were 
identified\footnote{For example, there was a missing a square root in going from a variance to a standard deviation.}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum 
estimates after cross-multiplying independent `even' and `odd' visibility measurements (which is done principally to avoid a noise bias). In \citetalias{ali_et_al2015}, a factor of $2$ was mistakenly used.
\item $N_{lst}$: The number of independent LST bins that go into a power spectrum estimation. The sensitivity scales as the square root 
because we integrate incoherently over time. For PAPER-64, $N_{lst} = 8$.
\item $N_{seps}$: The number of baseline separation types averaged incoherently in a final power spectrum estimate. For the 
analysis in this paper, we only use one type of baseline, hence $N_{seps}=1$. The updated limits in Kolopanis et al. (\textit{submitted})
use three separation types.
\item $t_{int}$: Length of an independent integration of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a 
fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time $t_{i}$
using the following:
\begin{equation}
t_{int} = t_{i} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{i}=43$ seconds, $t_{int}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is 
taken over all fringe-rates. For PAPER-64, this number is $t_{int} = 3857$\,s. 
\item $N_{days}$: The total number of days of data analyzed. In \citetalias{ali_et_al2015}, this number was set to $135$. However, because we 
divide our data in half (to form `even' and `odd' datasets), this number should be reduced by a factor of $2$. Additionally, 
because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a more realistic value of 
$N_{days}$ as:
\begin{equation}
 \frac{1}{N_{days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ is over LST (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{days}$ is $\sim34$ 
days.
\item $N_{bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In \citetalias{ali_et_al2015}, this number was 
the total number of $30$\,m East/West baselines used in the analysis. However, using the total number of baselines neglects 
the fact that baselines are averaged into groups (see Section \ref{sec:Boot}) for computational speed-up when cross-multiplying data. Our revised estimate for the parameter is:
\begin{equation}
N_{bls} = \frac{N_{bls\_total}}{N_{gps}}\sqrt{N_{gps}^{2}-N_{gps}},
\end{equation}
\noindent where, as with the \citetalias{ali_et_al2015} analysis, $N_{gps} = 5$. Each baseline group averages down linearly as the number of baselines 
entering the group ($N_{bls\_total}/N_{gps}$) and then as the square root of the number of cross-multiplied pairs ($\sqrt{N_{gps}^{2} - 
N_{gps}}$). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{bls} \sim 46$ instead 
of $51$. 
\item $N_{pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{pols}=2$.
\end{itemize}

@@

In Section \ref{sec:BiasOverview} we highlighted some common sources of bias that can show up as power spectrum 
detections and imitate EoR. We discussed the importance of using jackknife and null tests for instilling confidence in an EoR 
detection, as well as identifying other sources of biases. Here we demonstrate methods used by PAPER-64 to mitigate 
foreground and noise bias and we perform null tests in order to characterize the stability and implications of our results.

@@

\noindent Because of the nature of the lower triangular matrix, this window function has the property of preventing the leakage 
of foreground power from low $k$ to high $k$ modes. Specifically, we order the elements in $\textbf{F}$ in such a way so that 
power can leak from high $k$ modes to low $k$ modes, but not vice versa. Since most foreground power shows up at low 
$k$'s, this method ensures a window function that retains clean, noise-dominated measurements while minimizing the 
contamination of foreground bias. This tailored window function was used in the \citetalias{ali_et_al2015} analysis, however for this paper, we use $\textbf{M} \propto \textbf{I}$ for simplicity.

In addition to mitigating foreground bias at high $k$'s, two other sources of bias that we actively suppress in the PAPER-64 
analysis are noise bias associated with the squaring of thermal noise and noise bias from crosstalk. In order to avoid the 
former, we filter out certain cross-multiplications when forming $\widehat{q}$ in Equation \eqref{eq:qhat}. Namely, the PAPER-64 
dataset is divided into two halves: even Julian dates and odd Julian dates. Our data vectors are then $\textbf{x}_{even, 1}$ for 
the `even' dataset and baseline group $1$, $\textbf{x}_{odd, 1}$ for the `odd' dataset and baseline group $1$, etc. We only form 
$\widehat{q}$ when the two copies of $\textbf{x}$ come from different groups and baselines, never multiplying `even' with `even', for 
example, in order to prevent the squaring of the same thermal noise. We recognize that this method results in a hit to our sensitivity. Sensitivity can be gained, for example, by only dividing up the dataset once (either along time or baselines), or by creating more groups (more cross-multiplications), but in this paper we only attempt to revise the \citetalias{ali_et_al2015} analysis (which uses the same groupings), not produce the most sensitive limits of PAPER-64.

@@

As shown in Figure \ref{fig:ps1_data}, our re-analysis of the PAPER-64 power spectrum shows biases above the predicted noise level, particularly at low $k$ values. As discussed in Section \ref{sec:BiasTypes}, this bias is
most likely attributable to foreground leakage. The cause for biases at higher $k$ values is more difficult to pinpoint. Here we discuss three null tests performed on PAPER-64 data that aim to isolate systematics in the data and verify 
that they are not attributable to EoR. We build off of the framework in Section \ref{sec:JackknifeOverview}, which illuminated how jackknife tests can tease out excess variances along different axes of a dataset.

@@

\begin{itemize}
%\item Original Data (black): This is identical to the unweighted, revised PAPER-64 power spectrum in Figure \ref{fig:ps1_data} (one baseline type only) with weighting matrix $\textbf{I}$. There are clear detections at low $k$'s.
\item Baselines: We split our dataset into two halves, where each contains half of the total baselines used in the 
analysis. No baselines are repeated between the two datasets.
\item Sidereal Hour: We split our dataset into two halves based on LST, namely the first half (LSTs 0.5-4.6 hours) and second half (LSTs 
4.6-8.6 hours).
\item Day: We split our dataset into even and odd Julian days. We form power spectra for each separately, allowing the cross-multiplication of `even' with `even', for example, for this null test only. If the same sky signal is in both the `even' and `odd' datasets, we expect it to cancel out. %However, allowing these cross-multiplications can in practice introduce a low level noise bias... 
\end{itemize}

\begin{figure*}
	\centering
	\includegraphics[width=0.47\textwidth]{plots/null_bls.png}
	\includegraphics[width=0.47\textwidth]{plots/null_lsts.png}
	\includegraphics[width=0.47\textwidth]{plots/null_eo.png}
	\caption{Power spectrum results (with $2\sigma$ errors) for three null tests, where a jackknife is taken along the baseline axis (top left), LST axis (top right), and even/odd Julian date axis (bottom). The results shown are unweighted (no signal loss), where the power spectrum values plotted are computed from the difference between two power spectra produced on either side of the jackknife axis. The gray shaded region in each plot is the estimated $2\sigma$ theoretical noise limit given the parameters of each test. We find that there are no significant systematics that differ between the tests, though there are low level biases at high-$k$ values in both the LST test and even/odd test. All three tests, whether biased or not, have an extra noise bias that is strongly $k$-dependent. Residual foregrounds dominate at $k$-values near the horizon ($k\sim\pm 0.1$ $h$Mpc$^{-1}$), which is not surprising given that foregrounds are brightest in this region post-delay filtering.}
	\label{fig:null}
\end{figure*}

@@

We find that the baselines test is fully consistent with noise, as all error bars pass through zero. Therefore, we conclude that particular baselines are not to be blamed for the bias in our dataset. For the LST test, $90\%$ of all the error bars pass through zero, which is slightly lower than the expected $95\%$. This indicates a low level bias that is LST-dependent, and likely caused by residual foregrounds that vary in LST. In other words, the two jackknife halves differ by an amount greater than thermal noise and do not completely cancel out the sky signal when differencing. The even/odd test shows $81\%$ of all error bars being consistent with zero, again indicating the presence of a low level bias. The cause of this particular systematic is not obvious, and requires further investigation that we do not carry out in this paper.

Finally, we note that all three null tests exhibit error bars that are strongly $k$-dependent. The errors are largest just outside the horizon ($k\sim\pm 0.1$ $h$Mpc$^{-1}$) which is expected since foregrounds are brightest in this region post-delay filtering. We also note that all three tests, whether biased or unbiased, result in power spectrum errors that are slightly larger than the estimated thermal noise at high-$k$. This indicates an additive noise component that is increasing our errors (while not affecting whether they pass through zero or not).

There are a couple possible explanations for this. Although we expect each cross-multiplication that is used in power spectrum estimation to have independent noise, there is still the possibility of a noise-foreground coupling term that can introduce power. This is because cross-multiplications produce four additive terms - a signal-squared term (where signal includes both foregrounds and EoR), two cross-terms between the signal and noise, and one noise-only term. When differencing two power spectra, we expect the signal-only term to subtract out for a `passing' null test, and we expect the noise-only terms to be consistent with thermal noise. The cross-terms, however, can introduce a noise bias that would vary with foreground strength, therefore being consistent with the shape of our results. That being said, an alternate explanation for the large error bars seen at high $k$'s could be that our estimate of the thermal noise is too low, as all three tests seem to imply a larger, but consistent, estimate.

@@

\begin{itemize}
\item Substantial signal loss can result when weighting data using empirically estimated covariances (Section 
\ref{sec:SiglossOverview}). Loss of the 21\,cm signal is especially significant the fewer number of independent modes that
exist in the data. Hence, there exists a trade-off between sensitivity driven 
time-averaging techniques such as fringe-rate filtering and signal loss when using empirically estimated covariances. 
\item Signal injection and recovery simulations can be used to quantify signal loss (Section \ref{sec:Sigloss}). However, a 
signal-only simulation (i.e. comparing a uniformly weighted vs. weighted power spectrum of EoR only) can underestimate loss by 
failing to account for correlations between the data and signal (Section \ref{sec:siglossmethod}).
\item Errors that are estimated via bootstrapping can be underestimated if samples in the dataset are significantly correlated 
(Section \ref{sec:ErrorOverview}). If the number of independent samples in a dataset is well-determined, bootstrapping is a 
simple and accurate way of estimating errors.
\item Meaningful null tests are vital to validate an EoR detection (Section \ref{sec:JackknifeOverview}). Similarly, performing 
jackknife tests along multiple axes of a dataset is necessary for confidence in an EoR detection and can also be used to tease 
out systematics.
\end{itemize}

As a consequence of our investigations, we have also revised power spectrum results from PAPER-64. In Kolopanis et al., (\textit{submitted}), we quote a best $2\sigma$ power spectrum upper limit of $(218.7$ mK$)^{2}$ at $k=0.3$ $h$Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK. The reasons for a previously underestimated limit and 
ways in which our new analysis differs can be summarized by the following:

\begin{itemize}
\item Signal loss, previously found to be $<2\%$ in \citetalias{ali_et_al2015}, was under-estimated by a factor of $\sim$$1000$ for inverse 
covariance weighting. For our new analysis, we use a regularized covariance weighting method to minimize loss (Section 
\ref{sec:Weight}). However, because our revised weighting method is not as aggressive as the former, our results are still a 
factor of $\sim$$10$ in mK higher than previous limits. Under-estimated signal loss therefore represents the bulk of our revision. We 
note that PAPER's analysis is not the first to under-estimate loss; results from the GMRT (\citealt{paciga_et_al2013}) were also 
revised from new signal loss calculations associated with their singular value decomposition foreground filter.
\item Power spectrum errors, originally computed by bootstrapping, were under-estimated for the data by a factor of $\sim$$2$ in mK due to oversampling data whose effective number of independent samples was reduced from fringe-rate filtering (Section \ref{sec:Boot}). 
\item Several factors used in an analytic expression to predict the noise-level in PAPER-64 data were revised, yielding a 
decrease in predicted sensitivity level by a factor of $\sim$$2$ in mK (Section \ref{sec:PSSense}). We have verified our revised 
prediction extensively using pure noise simulations. We note that our sensitivity prediction is revised by a factor less than our 
power spectrum result, implying that if taken at face value, the theoretical prediction for noise in \citetalias{ali_et_al2015} was too high for its data 
points.
\end{itemize}

The future of 21\,cm cosmology is exciting, as new experiments have sensitivities that expect to reach and surpass EoR levels, improved 
foreground mitigation/removal strategies are being developed, and simulations are being designed to better understand 
instruments. On the power spectrum analysis side, robust signal loss simulations, precise error calculations, and 
comprehensive jackknife tests will play critical roles in accurate 21\,cm results. With strong foundations being established now, it is safe to say that we can expect to learn much about reionization and our early Universe in the coming years.

>>>>

While the null test is useful for testing noise properties and the uniformity of a dataset, jackknives are useful in pinpointing 
which data subsets are contaminated by biases and which are not; in our toy model we see that the bias exists only in $
\textbf{x}_{1}$ (Figure \ref{fig:toy_bias2}). If foreground or noise biases exist in a dataset, jackknives can tease them out and 
provide insight into possible sources. For example, if jackknives along the time-axis reveal a bias present at a certain LST, a 
likely explanation would be excess foreground emission from a radio source in the sky at that time. A jackknife test involving 
data before and after the application of a fringe-rate filter can reveal whether crosstalk noise bias is successfully suppressed 
with the filter, or if similar-shaped detections in both power spectra suggest otherwise. There are many other jackknife axes of 
which we will not go into detail here, including baseline, frequency, and polarization. Ultimately, an EoR detection should persist 
through them all and a clean measurement should exhibit noise-like null spectra.

@@

where $\nu$ are frequencies in GHz. We use a receiver temperature of $144$\,K, yielding $T_{sys} = 431$\,K at $150$\,MHz. 
This is lower than the $T_{sys}$ of $500$\,K used in \citetalias{ali_et_al2015} because of several small mis-calculation errors that were 
identified\footnote{For example, there was a missing a square root in going from a variance to a standard deviation.}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum 
estimates after cross-multiplying independent `even' and `odd' visibility measurements (which is done principally to avoid a noise bias). In \citetalias{ali_et_al2015}, a factor of $2$ was mistakenly used.
\item $N_{lst}$: The number of independent LST bins that go into a power spectrum estimation. The sensitivity scales as the square root 
because we integrate incoherently over time. For PAPER-64, $N_{lst} = 8$.
\item $N_{seps}$: The number of baseline separation types averaged incoherently in a final power spectrum estimate. For the 
analysis in this paper, we only use one type of baseline, hence $N_{seps}=1$. The updated limits in Kolopanis et al. (\textit{submitted})
use three separation types.
\item $t_{int}$: Length of an independent integration of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a 
fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time $t_{i}$
using the following:
\begin{equation}
t_{int} = t_{i} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{i}=43$ seconds, $t_{int}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is 
taken over all fringe-rates. For PAPER-64, this number is $t_{int} = 3857$\,s. 
\item $N_{days}$: The total number of days of data analyzed. In \citetalias{ali_et_al2015}, this number was set to $135$. However, because we 
divide our data in half (to form `even' and `odd' datasets), this number should be reduced by a factor of $2$. Additionally, 
because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a more realistic value of 
$N_{days}$ as:
\begin{equation}
 \frac{1}{N_{days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ is over LST (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{days}$ is $\sim34$ 
days.
\item $N_{bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In \citetalias{ali_et_al2015}, this number was 
the total number of $30$\,m East/West baselines used in the analysis. However, using the total number of baselines neglects 
the fact that we average baselines into groups for computational speed-up when cross-multiplying data. Our revised estimate for the parameter is:
\begin{equation}
N_{bls} = \frac{N_{bls\_total}}{N_{gps}}\sqrt{N_{gps}^{2}-N_{gps}},
\end{equation}
\noindent where, as with the \citetalias{ali_et_al2015} analysis, $N_{gps} = 5$. Each baseline group averages down linearly as the number of baselines 
entering the group ($N_{bls\_total}/N_{gps}$) and then as the square root of the number of cross-multiplied pairs ($\sqrt{N_{gps}^{2} - 
N_{gps}}$). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{bls} \sim 46$ instead 
of $51$. 
\item $N_{pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{pols}=2$.
\end{itemize}

@@

In Section \ref{sec:BiasOverview} we highlighted some common sources of bias that can show up as power spectrum 
detections and imitate an EoR signal. We discussed the importance of using jackknife and null tests for instilling confidence in an EoR 
detection, as well as identifying other sources of biases. Here we demonstrate methods used by PAPER-64 to mitigate 
foreground and noise bias and we perform null tests in order to characterize the stability and implications of our results.

@@

\noindent Because of the nature of the lower triangular matrix, this window function has the property of preventing the leakage 
of foreground power from low-$k$ to high-$k$ modes. Specifically, we order the elements in $\textbf{F}$ in such a way so that 
power can leak from high-$k$ modes to low-$k$ modes, but not vice versa. Since most foreground power shows up at low-$k$'s, this method ensures a window function that retains clean, noise-dominated measurements while minimizing the 
contamination of foreground bias. This tailored window function was used in the \citetalias{ali_et_al2015} analysis, however throughout this paper, we use $\textbf{M} \propto \textbf{I}$ for simplicity.

In addition to mitigating foreground bias at high-$k$'s, two other sources of bias that we actively suppress in the PAPER-64 
analysis are noise bias associated with the squaring of thermal noise and noise bias from crosstalk. In order to avoid the 
former, we filter out certain cross-multiplications when forming $\widehat{q}$ in Equation \eqref{eq:qhat}. Namely, the PAPER-64 
dataset is divided into two halves: even Julian dates and odd Julian dates. Our data vectors are then $\textbf{x}_{even, 1}$ for 
the `even' dataset and baseline group $1$, $\textbf{x}_{odd, 1}$ for the `odd' dataset and baseline group $1$, etc. We only form 
$\widehat{q}$ when the two copies of $\textbf{x}$ come from different groups and baselines, never multiplying `even' with `even', for 
example, in order to prevent the squaring of the same thermal noise. We recognize that this method results in a hit to our sensitivity. Sensitivity can be gained, for example, by only dividing up the dataset once (either along time or baselines), or by creating more groups (more cross-multiplications), but in this paper we only attempt to revise the \citetalias{ali_et_al2015} analysis (which uses the same groupings), not produce the most sensitive limits of PAPER-64.

@@

As shown in Figure \ref{fig:ps1_data}, our re-analysis of the PAPER-64 power spectrum shows biases above the predicted noise level, particularly at low-$k$ values. As discussed in Section \ref{sec:BiasTypes}, this bias is
most likely attributable to foreground leakage. The cause for biases at higher $k$-values is more difficult to pinpoint. Here we discuss three null tests performed on PAPER-64 data that aim to isolate systematics in the data and verify 
that they are not attributable to EoR. We build off of the framework in Section \ref{sec:JackknifeOverview}, which illuminated how jackknife tests can tease out excess variances along different axes of a dataset.

@@

\begin{itemize}
%\item Original Data (black): This is identical to the unweighted, revised PAPER-64 power spectrum in Figure \ref{fig:ps1_data} (one baseline type only) with weighting matrix $\textbf{I}$. There are clear detections at low $k$'s.
\item Baselines: We split our dataset into two halves, where each contains half of the total baselines used in the 
analysis. No baselines are repeated between the two datasets.
\item Sidereal Hour: We split our dataset into two halves based on LST, namely the first half (LSTs 0.5-4.6 hours) and second half (LSTs 
4.6-8.6 hours).
\item Day: We split our dataset into even and odd Julian dates. We form power spectra for each separately, allowing the cross-multiplication of `even' with `even', for example, for this null test only. If the same sky signal is in both the `even' and `odd' datasets, we expect it to cancel out. %However, allowing these cross-multiplications can in practice introduce a low level noise bias... 
\end{itemize}

\begin{figure*}
	\centering
	\includegraphics[width=0.47\textwidth]{plots/null_bls.png}
	\includegraphics[width=0.47\textwidth]{plots/null_lsts.png}
	\includegraphics[width=0.47\textwidth]{plots/null_eo.png}
	\caption{Differenced power spectrum results (with $2\sigma$ errors) for three null tests, where a jackknife is taken along the baseline axis (top left), LST axis (top right), and even/odd Julian date axis (bottom). The results shown are unweighted (no signal loss), where the power spectrum values plotted are computed from the difference between two power spectra produced on either side of the jackknife axis. The gray shaded region in each plot is the estimated $2\sigma$ theoretical noise limit given the parameters of each test. We find that there are no significant systematics that differ between the tests, though there are low level biases at high-$k$ values in both the LST test and even/odd test. All three tests, whether biased or not, have an extra noise bias that is strongly $k$-dependent. Residual foregrounds dominate at $k$-values near the horizon ($k\sim\pm 0.1$ $h$Mpc$^{-1}$), which is not surprising given that foregrounds are brightest in this region post-delay filtering.}
	\label{fig:null}
\end{figure*}

@@

We find that the baselines test is fully consistent with noise, as all error bars pass through zero. Therefore, we conclude that particular baselines are not to be blamed for the bias in our dataset. For the LST test, $90\%$ of all the error bars pass through zero, which is slightly lower than the expected $95\%$. This indicates a low level bias that is LST-dependent, and likely caused by residual foregrounds that vary in LST. In other words, the two jackknife halves differ by an amount greater than thermal noise and do not completely cancel out the sky signal when differencing. The even/odd test shows $81\%$ of all error bars being consistent with zero, again indicating the presence of a low level bias. The cause of this particular systematic is not obvious, and requires further investigation that we do not carry out in this paper. Most importantly, a clean detection of EoR would have passed all three null tests, which is clearly not what we see.

Finally, we note that all three null tests exhibit error bars that are strongly $k$-dependent. The errors are largest just outside the horizon ($k\sim\pm 0.1$ $h$Mpc$^{-1}$) which is expected since foregrounds are brightest in this region post-delay filtering. We also note that all three tests, whether biased or unbiased, result in power spectrum errors that are slightly larger than the estimated thermal noise (gray band) at high-$k$. This indicates an additive noise component that is increasing our errors (while not affecting whether they pass through zero or not).

There are a couple possible explanations for this. Although we expect each cross-multiplication that is used in power spectrum estimation to have independent noise, there is still the possibility of a noise-foreground coupling term that can introduce power. This is because cross-multiplications produce four additive terms - a signal-squared term (where signal includes both foregrounds and EoR), two cross-terms between the signal and noise, and one noise-only term. When differencing two power spectra, we expect the signal-only term to subtract out for a `passing' null test, and we expect the noise-only terms to be consistent with thermal noise. The cross-terms, however, can introduce a noise bias that would vary with foreground strength, therefore being consistent with the shape of our results. That being said, an alternate explanation for the large error bars seen at high $k$'s could simply be that our estimate of the thermal noise is too low, as all three tests seem to imply a larger, but consistent, estimate.

@@

\begin{itemize}
\item Substantial signal loss can result when weighting data using empirically estimated covariances (Section 
\ref{sec:SiglossOverview}). Loss of the 21\,cm signal is especially significant the fewer number of independent modes that
exist in the data. Hence, there exists a trade-off between sensitivity driven 
time-averaging techniques such as fringe-rate filtering and signal loss when using empirically estimated covariances. 
\item Signal injection and recovery simulations can be used to quantify signal loss (Section \ref{sec:Sigloss}). However, a 
signal-only simulation (i.e. comparing a uniformly weighted vs. weighted power spectrum of EoR only) can under-estimate loss by 
failing to account for correlations between the data and signal (Section \ref{sec:siglossmethod}).
\item Errors that are estimated via bootstrapping can be under-estimated if samples in the dataset are significantly correlated 
(Section \ref{sec:ErrorOverview}). However, if the number of independent samples in a dataset is well-determined, bootstrapping is a 
simple and accurate way of estimating errors.
\item Meaningful null tests are vital to validate an EoR detection (Section \ref{sec:JackknifeOverview}). Similarly, performing 
jackknife tests along multiple axes of a dataset is necessary for confidence in an EoR detection and can also be used to tease 
out systematics.
\end{itemize}

As a consequence of our investigations, we have also revised power spectrum results from PAPER-64. In Kolopanis et al. (\textit{submitted}), we quote a best $2\sigma$ power spectrum upper limit of $(240.0$ mK$)^{2}$ at $k=0.3$ $h$Mpc$^{-1}$ and $z=8.37$, 
a higher limit than \citetalias{ali_et_al2015} by a factor of $\sim$$10$ in mK. The reasons for a previously under-estimated limit and 
ways in which our new analysis differs can be summarized by the following:

\begin{itemize}
\item Signal loss, previously found to be $<2\%$ in \citetalias{ali_et_al2015}, was under-estimated by a factor of $>$$1000$ for empirically estimated inverse 
covariance weighting. For our new analysis, we use a regularized covariance weighting method to minimize loss (Section 
\ref{sec:Weight}). However, because our revised weighting method is not as aggressive as the former, our results are still a 
factor of $\sim10$ in mK higher than previous limits. Under-estimated signal loss therefore represents the bulk of our revision. We 
note that PAPER's analysis is not the first to under-estimate loss; results from the GMRT (\citealt{paciga_et_al2013}) were also 
revised from new signal loss calculations associated with their singular value decomposition foreground filter.
\item Power spectrum errors, originally computed by bootstrapping, were under-estimated for the data by a factor of $\sim2$ in mK due to oversampling data whose effective number of independent samples was reduced from fringe-rate filtering (Section \ref{sec:Boot}). 
\item Several factors used in an analytic expression to predict the noise-level in PAPER-64 data were revised, yielding a 
decrease in predicted sensitivity level by a factor of $\sim2$ in mK (Section \ref{sec:PSSense}). We have verified our revised 
prediction extensively using pure noise simulations. We note that our sensitivity prediction is revised by a factor less than our 
power spectrum result, implying that if taken at face value, the theoretical prediction for noise in \citetalias{ali_et_al2015} was too high for its data 
points.
\end{itemize}

The future of 21\,cm cosmology is exciting, as new experiments have sensitivities that expect to reach and surpass EoR levels, improved 
foreground mitigation and removal strategies are being developed, and simulations are being designed to better understand 
instruments. On the power spectrum analysis side, robust signal loss simulations, precise error calculations, and 
comprehensive jackknife tests will play critical roles in accurate 21\,cm results. With strong foundations being established now, it is safe to say that we can expect to learn much about reionization and our early Universe in the coming years.
